---
title: "NLPWorkshop_Hills"
author: "Thomas T. Hills"
output: html_document
date: "2024-09-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r installlibraries}
library(readr)
library(rstudioapi)
library(tidyverse)
library(tidytext)
library(wordcloud)
library(wordcloud2)
library(quanteda)
library(zoo) # rollapply
library(tm)
library(lsa)
library(LDAvis)
library(topicmodels)
```

```{r datahygeine}
rm(list=ls())
```

# Read in txt files

```{r}
# select input directory with text files
input_directory <- selectDirectory()

# Alternatively, you can set the directory
# input_directory = './NLPWorkshopData/text_files'   # set sets the path directory to our text files

# get list of files in directory
filesList<- list.files(input_directory)

filesList

documentVector <- c()

for(i in 1:length(filesList)){
 # classData <- read.csv(paste(input_directory, "/",fileToRead, sep=""), header = T)
  rd <- read_lines(file=paste(input_directory, "/", filesList[i],sep=""))
  rd2 <- paste(rd, sep =" ", collapse = " ")
  documentVector <- c(documentVector, rd2)
}

# make into a data frame
dv <- tibble(docnum = 1:length(documentVector), docName = filesList, text = documentVector)

# Note: we could read in a csv file and label it as above to achieve the same thing if we had a data file with a document in each row
```

# Preprocessing

```{r preprocessing}
# Change encoding to convert non-ascii characters to ascii (ü -> u)
# you'll want to do this to norms to if needed
  dv$text <- iconv(dv$text, from = "UTF-8", to = "ASCII", sub = "")
    #### clean data file
  dv$text <- iconv(dv$text, "latin1", "ASCII", sub="") # remove non-asci characters
  dv$text <- gsub("[[:punct:]]","",dv$text) # remove all puncutation
  dv$text <- gsub('[[:digit:]]+', "", dv$text) # removes digits
  dv$text <- tolower(dv$text) # make it all lower case
 
```

# Word counting

## unnest tokens

```{r unnest}
dvun <- dv %>% unnest_tokens(word, text)
```

## Finding words

```{r searchForWords}
"cheshire" %in% dvun$word
"hatter" %in% dvun$word
"snake" %in% dvun$word
"caterpillar" %in% dvun$word
```

## Word frequency

```{r countwords}
dvun %>% dplyr::count(word, sort=TRUE) 
FreqCount <- dvun %>% count(word, sort=TRUE) 

dvun %>% dplyr::count(word, sort=TRUE) %>% dplyr::arrange(-dplyr::row_number())

```

## Remove stopwords

```{r stopwords}
data('stop_words') 

dvNoStops <- dvun %>% anti_join(subset(stop_words, lexicon == "snowball")) # this limits stopwords to snowball stopwords

dvNoStops %>% dplyr::count(word, sort = TRUE)
```

## Stemming

```{r stemming}
text <- "love loving lovingly Loved lover lovely love"
text <- tibble(text=text)
text <- text %>% unnest_tokens(word, text)

text %>%
  mutate(word = wordStem(word))

dvNoStopsStemmed <- dvNoStops %>%
  mutate(word_stem = wordStem(word))

```

## Wordclouds

```{r wordclouds}
dvun %>%
  dplyr::count(word, sort = TRUE) %>%
  with(wordcloud(word, n, max.words = 10, colors=brewer.pal(8, "Dark2")))

dvNoStops %>%
  anti_join(stop_words) %>%
  dplyr::count(word, sort = TRUE) %>%
  with(wordcloud(word, n, max.words = 20, colors=brewer.pal(8, "Dark2")))

```

## Concordances

```{r concordances}
library(quanteda)
dvt <- quanteda::tokens(dv$text,
            remove_punct = T, remove_symbols = T)
quanteda::kwic(dvt, pattern = "queen", window = 3)
# to find a phrase
quanteda::kwic(dvt, pattern =  phrase("mock turtle"), window = 3)
quanteda::kwic(dvt, pattern =  phrase("freedom"), window = 3)

```

## Bigrams

```{r bigrams}
dvBigrams <- dv %>%
  unnest_tokens(bigram, text, token = "ngrams", n = 2) # n = 2 means we want 2-grams

bigramsSplit <- dvBigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigramsSplit %>% # take out all bigrams with stopwords
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

bigrams_united <- bigrams_filtered %>% # reunite the bigrams
  unite(bigram, word1, word2, sep = " ")

bigram_counts <- bigrams_united %>%
  dplyr::group_by(docName) %>%
  dplyr::count(bigram, sort = TRUE)


# First, filter out bigrams that occur only once
bisgt1 <- bigram_counts %>%
  filter(n > 1)

df_top10 <- bisgt1 %>%
  group_by(docName) %>%
  top_n(10, n) %>%
  ungroup()

df_top10 %>%
  filter(bigram != "NA NA") %>%
  #mutate(bigram= factor(bigram, levels = rev(unique(bigram)))) %>% 
  group_by(docName) %>% 
  top_n(10) %>% 
  ungroup() %>%
  ggplot(aes(reorder(bigram,n), n, fill = docName)) +
  geom_col(show.legend = FALSE) +
  labs(x = NULL, y = "count") +
  facet_wrap(~docName, ncol = 2, scales = "free") +
  coord_flip()

```

# Regular expression to combine names

```{r}
# gsub finds one pattern and replaces it with another
text_mock_turtle <- gsub("mock turtle", "mock_turtle", dv$text)

dvt <- quanteda::tokens(dv$text, remove_punct = T, remove_symbols = T)
dvsub <- quanteda::tokens(text_mock_turtle, remove_punct = T, remove_symbols = T)
quanteda::kwic(dvt, pattern = phrase("mock turtle"), window = 3)
quanteda::kwic(dvsub, pattern = phrase("mock_turtle"), window = 3)
quanteda::kwic(dvsub, pattern = phrase("mock turtle"), window = 3)

```

# Feature analysis (sentiment)

```{r sentiment_analysis}
# get norms

vad <- read.csv("NLPWorkshopData/norms/vadNorms.csv")
vad <- tibble(vad)

# left_join with text file
vaddvun <- dvun %>% left_join(vad)

# what are the most negative words 
vaddvun %>% arrange(valence)

# what are the most positive words
vaddvun %>% arrange(desc(valence))

# what's the mean valence for all documents?
vaddvun %>% summarise(mean(valence, na.rm=T))

vaddvun %>%  group_by(docName) %>% summarise(valence = mean(valence, na.rm=T))
```

## Feature analysis (multiple documents and output) for all norms in norm folder

```{r featureMultiple}



###### first we'll read in the norms: we'll compute these for each text
###### if you want to add new norms, just add them to the `norms` folder (properly formatted with `word` and `ratingName`)

normsFiles <- list.files('NLPWorkshopData/norms/')
length(normsFiles)
normF <- NULL
for(i in 1:length(normsFiles)){
  normF[[i]] <- read.csv(paste("NLPWorkshopData/norms/", normsFiles[i],sep=""))
}

# join all norms
docRatings <- dvun 
for(i in 1:length(normF)) {
   docRatings <- docRatings %>% left_join(normF[[i]], by = "word")
}

# Loop through all docs and compute stats for each 
# Because I don't trust group_by I use a for loop
docs <- unique(docRatings$docName)
dataout <- c()
for(i in 1:length(docs)){
  buf <- subset(docRatings, docName == docs[i])
  tokenn <- nrow(buf) 
  typesn <- length(unique(buf$word))
  TTR <- typesn/tokenn
  # what's the mean of the values for this document
  values <- names(docRatings)[4:length(names(docRatings))]
  dataVector <- buf %>% summarise_at(values, mean,na.rm=T)
  dataVector <- data.frame(filename = docs[i], dataVector, tokens = tokenn, types=typesn, TTR=TTR)
  dataout <- rbind(dataout, dataVector)
}

write.table(dataout, "output.txt", quote = FALSE, row.names = FALSE)


with(dataout, plot(ageOfAcquisition, concreteness, ylim = c(2.2, 2.6), xlim = c(4.6, 6)))
with(dataout, text(ageOfAcquisition, concreteness+.03, labels=filename, cex = .5))

```

### Moving Window

We using a *moving window* to see when things turn up or down. A *window* simply refers to a set of data we will look at together as a kind of chunk or bin. But with a moving window, we move the window over one item, and compute again, and do this repeatedly across the document one word at a time. This smooths out the data by moving the window--one word at a time-- over the text.

`rollapply` makes this incredibly easy to do as. It simply takes a vector and computes a value for each row in the vector based on the values above and below that row. It has a variety of parameters to handle the two ends of the document as well as other decisions. You can learn more about these by looking at `?rollapply`.

```{r}

# lets focus on a specific document
alice <- docRatings[docRatings$docName=="AlicesAdventuresInWonderland_symbols.txt",]
# make line numbers
alice <- tibble(line=1:nrow(alice), alice)
# set windowsize 
windowSize = 500

windowValue <- rollapply(alice$valence, windowSize, FUN = "mean",align="center", na.rm=T, partial = TRUE)

rollingAlice <- alice %>% mutate(windowValue)

with(rollingAlice, plot(line, windowValue, cex = .2, pch=16, xlab = "Word in Alice", ylab = "Valence"))

```

Try adjusting the `windowSize` in the above code to see what it does. I've created a simple tool called [textsight](https://warwick.ac.uk/fac/sci/psych/people/thills/thills/textsight/) where you can investigate this further. You can also upload your own documents there and mouse-over to read individual parts of the text and quickly visualize documents, like reviewer's comments to determine which reviewer is in fact the most evil. Using other tools here you could probably determine when you're dealing with the same reviewer.


# Semantics


## LSA

```{r}

# Clean the text a bit before making the corpus
for(i in 1:nrow(dv)){
  dv$text[i] <- iconv(dv$text[i], from = "UTF-8", to = "ASCII", sub = "")
  # add in lines to remove numbers
}

# try again with corpus
corpus <- Corpus(VectorSource(dv$text))

# Preprocess the text (gives warnings, but they are not relevant here)
corpus <- tm_map(corpus, content_transformer(tolower))
corpus <- tm_map(corpus, removePunctuation)
corpus <- tm_map(corpus, removeNumbers)
corpus <- tm_map(corpus, removeWords, stopwords("en"))
corpus <- tm_map(corpus, stripWhitespace)

# Create a Term-Document Matrix
tdm <- TermDocumentMatrix(corpus)

# Inspect the Term-Document Matrix
inspect(tdm)

dim(tdm)

tdm$dimnames$Docs <- dv$docName

termcount <-apply(tdm,1,sum)
doccount <- apply(tdm, 2, sum)

# this needs to be a matrix (which isn't made automatically above)
TDM <- as.matrix(tdm)

TDM2 <- lw_tf(TDM) * gw_idf(TDM)

LSAout <- lsa(TDM2, dims=dimcalc_share())
#dimcalc_share() computes a 'recommended' number of dimensions from the data. You could choose this by hand.  

head(LSAout$dk)  # the document matrix
head(LSAout$tk)  # the term matrix
head(LSAout$sk)  # the diagonal matrix

myDocs <- rownames(LSAout$dk) 
head(myDocs)
myTerms <- rownames(LSAout$tk) 
head(myTerms)

termd = (LSAout$sk * t(LSAout$tk))
termd[1:5, 1:5]
#termd = (LSAout$tk * LSAout$sk)
#termd[1:5, 1:5]

cosine_terms <- cosine(termd) # this takes some time 

# see top 10
cosine_terms[1:10,1:10]

wtf <- "america"
cosine_terms[wtf,order(cosine_terms[wtf,], decreasing=T)][1:10]

# the same for documents
docd = (LSAout$sk * t(LSAout$dk))

cosine_docs <- cosine(docd)

cosine_docs[1:5, 1:5]

wtf <- "1957-Eisenhower.txt"
cosine_docs[wtf,order(cosine_docs[wtf,], decreasing=T)][1:10]

```

Finally, we can visualize all of this, to see how similar documents are to one another in a visual way. This uses multi-dimensional scaling, which embeds many dimensions into a two dimensional space, allowing us to plot them.

```{r}
# MDS to visualize similarities
d <- dist(cosine_docs) # euclidean distances between the rows
fit <- cmdscale(d,eig=TRUE, k=2) # k is the number of dim
fit$points # view results

# plot solution 
x <- fit$points[,1]
y <- fit$points[,2]
plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2", 
     main="Metric MDS", type="n")
text(x, y, labels = row.names(cosine_docs), cex=.5)

```

Lovely.


# LDA  Topics Analysis

```{r LDA}

allFiles <- dvun 


# prepare for a document term matrix
word_counts <- allFiles %>%
  anti_join(stop_words) %>%
  dplyr::count(docName, word, sort = TRUE) 

# the method we use here expects to receive a count for each term in each document
# cast into a term-document-matrix --> a different method from before, but does the same thing
addresses_dtm <- word_counts %>%
  cast_dtm(docName, word, n)

addresses_dtm

k = 13 # Number of topics 

addresses_lda <- LDA(addresses_dtm, k = k, control = list(seed = 1, alpha = .1 ))

ad_topics <- tidy(addresses_lda, matrix = "beta") # Note that with k = k topics we have k * the number of terms in the dtm

# show most probable terms
terms(addresses_lda, k, threshold= 0)

ad_top_terms <- ad_topics %>%
  group_by(topic) %>%
  top_n(10, beta) %>% # Take Top 10 terms
  ungroup() %>%
  arrange(topic, -beta)


ad_top_terms %>%
  mutate(term = reorder_within(term, beta, topic)) %>% # order within topics
  ggplot(aes(term, beta, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip() +
  scale_x_reordered()



## compute LIFT

ad_lift <- ad_topics %>% group_by(term) %>%
  mutate(term_topic_prob = beta / sum(beta,na.rm=T))
  # we divide by the sum of the Betas for the term across topics
  

ad_top_lift <- ad_lift %>%
  group_by(topic) %>%
  top_n(10, term_topic_prob) %>%
  sample_n(10) %>%
  ungroup() %>%
  arrange(topic, -term_topic_prob)


ad_top_lift %>% 
  mutate(term = reorder_within(term, term_topic_prob, topic)) %>%
  ggplot(aes(term, term_topic_prob, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip() +
  scale_x_reordered()

# Relevance (from Sievert and Shirley)

lambda = .6 # this tunes between raw beta (=1) and lift (=0)

ad_rel <- ad_topics %>% group_by(term) %>%
  mutate(relevance = lambda* log(beta) + (1-lambda)* log(beta / sum(beta,na.rm=T)) )
  

ad_top_rel <- ad_rel %>%
  group_by(topic) %>%
  top_n(10, relevance) %>%
  sample_n(10) %>%
  ungroup() %>%
  arrange(topic, -relevance)

ad_top_rel2 <- ad_rel %>%
  group_by(topic) %>%
  top_n(10, relevance) %>%
  ungroup() %>%
  arrange(topic, -relevance)


ad_top_rel %>% 
  mutate(term = reorder_within(term, relevance, topic)) %>%
  ggplot(aes(term, relevance, fill = factor(topic))) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ topic, scales = "free") +
  coord_flip() +
  scale_x_reordered()


# this allows you to look at topic 1 lift values
ad_top_lift %>% filter(topic == 1)
ad_top_lift %>% filter(topic == 2)

ldaOut <- addresses_lda

# Gamma gives the topic probability for each document
adGamm <- tidy(ldaOut, matrix = "gamma")
adGamm

adGamm <- adGamm %>%
  separate(document, c("year", "president"), sep = "-", convert = TRUE)

adGamm

adGamm %>%
ggplot(aes(year, gamma, group = factor(topic))) +
        geom_smooth(aes(color = factor(topic)),se=F, span = .3)
        
ldaOut.topics <- as.matrix(topics(ldaOut))


# write.csv(ldaOut.topics,file=paste("topic_model",k,"DocsToTopics.csv"))

ldaOut.terms <- as.matrix(terms(ldaOut,6))
ldaOut.terms[1:6,]


topicmodels2LDAvis <- function(x, ...){
  post <- topicmodels::posterior(x)
  if (ncol(post[["topics"]]) < 3) stop("The model must contain > 2 topics")
  mat <- x@wordassignments
  LDAvis::createJSON(
    phi = post[["terms"]], 
    theta = post[["topics"]],
    vocab = colnames(post[["terms"]]),
    doc.length = slam::row_sums(mat, na.rm = TRUE),
    term.frequency = slam::col_sums(mat, na.rm = TRUE)
  )
}

#
serVis(topicmodels2LDAvis(ldaOut))

dir_name  = paste('result_lda/LDA_k=', k, sep='')

#serVis(topicmodels2LDAvis(ldaOut), out.dir = dir_name, open.browser = TRUE) ## the output explains what location to find the visualization at.  You may need to copy and paste that into your web browser. It starts with `http://`.


```
